---
title: "Somerset ICB Demand Modelling Summary"
subtitle: "Using Quality and Outcomes Framework (QOF) Prevalence"
author: "Richard Blackwell"
format:
  revealjs:
    logo: HISW_Logo_RGB_Positive.png
    theme: [default, hisw.scss]
    embed-resources: true
    code-copy: false
    view-distance: 10
---

```{r}
#| echo: false
#| warning: false

library(tidyverse)
library(readxl)
library(plotly)
library(ggiraph)

# Ensure directory is correct
if(tolower(basename(getwd()))!="somerset_icb")
  setwd("./somerset_icb")

# Load input data ----
QOF <- FALSE

if(QOF){
  filename <- "./input/qof_input.xlsx"
  sim_filename <- "simulation_grouped_qof.csv"
  sim_zipfile <- "./output/qof/simulation_grouped_qof.zip"
} else {
  filename <- "./input/gbd_input.xlsx"
  sim_filename <- "simulation_grouped_gbd.csv"
  sim_zipfile <- "./output/gbd/simulation_grouped_gbd.zip"
}
  
popn_sheet <- "POPN"
popn_growth_sheet <- "GROWTH"
prev_sheet <- "PREV"
odds_ratio_sheet <- "OR"
conditions_sheet <- "COND"

# Load population data ----
df_popn <- readxl::read_xlsx(path = filename, sheet = popn_sheet)

# Load population growth data ----
df_popn_growth <- readxl::read_xlsx(path = filename, sheet = popn_growth_sheet)

# Load prevalence data ----
df_prev <- readxl::read_xlsx(path = filename, sheet = prev_sheet) %>%
  pivot_wider(id_cols = "IDX", names_from = "COND", values_from = "PREV", values_fill = 0)

# Load odds ratio data ----
df_or <- readxl::read_xlsx(path = filename, sheet = odds_ratio_sheet)

# Load condition name lookup ----
df_cond <- readxl::read_xlsx(path = filename, sheet = conditions_sheet)

# Load benchmark population ----
age_bands <- c(paste0(str_sub(paste0("0", seq(0, 89, 5)), -2), "_", str_sub(paste0("0", seq(4, 89, 5)), -2)), "90_PLUS", "90_PLUS")
names(age_bands) <- c(paste0(seq(0, 90, 5), "_", seq(4, 94, 5)), "95+")

df_benchmark_popn <- read.csv("./input/gp-reg-pat-prac-quin-age.csv") %>%
  filter(ORG_TYPE=="Comm Region" & SEX!="ALL" & AGE_GROUP_5!="ALL") %>% 
  mutate(IDX = paste0(str_sub(SEX, 1, 1), "|", unname(age_bands[AGE_GROUP_5])), SIZE = NUMBER_OF_PATIENTS, .keep = "none") %>%
  group_by(IDX) %>%
  summarise(SIZE = sum(SIZE)) %>%
  ungroup()

# Load grouped simulation data
df_sim <- read.csv(unzip(zipfile = sim_zipfile, files = sim_filename))

# Remove unzipped file
ret <- file.remove(sim_filename)
```


## Inputs

The inputs for the demand modelling consist of

- Target population
- Population growth estimate
- Prevalence estimate
- Multi-morbidity estimate

## Initial Population {.smaller}

```{r}
#| echo: false
#| warning: false

df_plot_data <- df_popn %>% 
  mutate(
    AREA = "Somerset ICB",
    GENDER = str_sub(IDX, 1, 1),
    AGE_BAND = str_sub(IDX, 3),
    AGE_BAND_LABEL = if_else(str_sub(IDX, 3)=="90_PLUS", "90+", gsub("_", "-", AGE_BAND)),
    SIZE, PCT = SIZE/sum(SIZE),
    LABEL = sprintf("<b>%s</b><br>%s<br>%s (%.1f%%)",
                    AREA, paste0(if_else(GENDER=="F", "Female", "Male"), " ", AGE_BAND_LABEL),
                    prettyNum(SIZE, digits = 1, big.mark = ","), PCT*100),
    .keep = "none") %>%
  bind_rows(df_benchmark_popn %>% 
              mutate(
                AREA = "England",
                GENDER = str_sub(IDX, 1, 1),
                AGE_BAND = str_sub(IDX, 3),
                AGE_BAND_LABEL = if_else(str_sub(IDX, 3)=="90_PLUS", "90+", gsub("_", "-", AGE_BAND)),
                SIZE, PCT = SIZE/sum(SIZE),
                LABEL = sprintf("<b>%s</b><br>%s<br>%s (%.1f%%)",
                                AREA, paste0(if_else(GENDER=="F", "Female", "Male"), " ", AGE_BAND_LABEL),
                                prettyNum(SIZE, digits = 1, big.mark = ","), PCT*100),
            .keep = "none"))

# Ensure consistent category order for age bands:
age_bands_cat <- unique(if_else(unname(age_bands)=="90_PLUS", "90+", gsub("_","-",unname(age_bands))))
tick_vals  <- pretty(c(-max(df_plot_data$PCT), max(df_plot_data$PCT)))
tick_text  <- scales::percent(abs(tick_vals))
chart_title <- "Population Pyramid of<br>NHS Somerset ICB vs England"


plt_popn <- plot_ly() %>%
  add_trace(
    data = df_plot_data %>% filter(AREA == "Somerset ICB"),
    x = ~if_else(GENDER=="M", PCT, -PCT),
    y = ~AGE_BAND_LABEL,
    text = ~LABEL,
    color = ~GENDER,
    colors = c("M" = "#293894", "F" = "#00abbd"),
    type = "bar",
    orientation = "h",
    textposition = "none",
    hovertemplate = "%{text}<extra></extra>"
  ) %>%
  add_trace(
    data = df_plot_data %>% filter(AREA == "England"),
    name = "England",
    x = ~if_else(GENDER=="M", PCT, -PCT),
    y = ~AGE_BAND_LABEL,
    text = ~LABEL,
    marker = list(color = "#ffffff64",
                  line = list(color = "#211752", width = 1.5)),
    type = "bar",
    orientation = "h",
    textposition = "none",
    hovertemplate = "%{text}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    barmode = "overlay",
    xaxis = list(
      title = "% of Population",
      tickmode = "array",
      tickvals = tick_vals,
      ticktext = tick_text,
      zeroline = TRUE
    ),
    yaxis = list(
      title = "Age Band",
      categoryorder = "array",
      categoryarray = age_bands_cat
    ),
    legend = list(orientation = "h", x = 0, y = -0.05),
    margin = list(l = 70, r = 30, t = 50, b = 40)
  )
```

:::: {.columns}

::: {.column style="width: 30%; font-size: 18px;"}

The initial population is the NHS Somerset Integrated Care Board population as at 2025 by gender and 5 year age band^[[NHS Digital Patients Registered at a GP Practice](https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice)]

From the population pyramid we can see that there is a lower proportion of under 50-year-olds in the Somerset population, and notably lower in the 20-49 year old working age population and conversely a higher proportion of over 50-year-olds.

:::

::: {.column width="70%"}

```{r}
#| fig.width: 4
#| fig.height: 6
#| out-width: 100%

plt_popn %>% config(responsive = TRUE)
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

::::

## Population Growth {.smaller}

```{r}
#| echo: false
#| warning: false

df_plot_data <- df_popn %>%
  left_join(df_popn_growth, by = "IDX") %>%
  mutate(
    YEAR,
    GENDER = str_sub(IDX, 1, 1),
    AGE_BAND = str_sub(IDX, 3),
    AGE_BAND_LABEL = if_else(str_sub(IDX, 3)=="90_PLUS", "90+", gsub("_", "-", AGE_BAND)),
    SIZE = SIZE * (1 + PCT_GROWTH),
    .keep = "none") %>%
  group_by(YEAR) %>%
  mutate(PCT = SIZE/sum(SIZE)) %>%
  ungroup() %>%
  mutate(
    LABEL = sprintf("<b>%s</b><br>%s<br>%s (%.1f%%)",
                    YEAR, paste0(if_else(GENDER=="F", "Female", "Male"), " ", AGE_BAND_LABEL),
                    prettyNum(SIZE, digits = 1, big.mark = ","), PCT*100))

chart_title <- sprintf("Population Pyramid Projection<br>for Somerset ICB")
  
# Ensure consistent category order for age bands:
# Age levels as above
tick_vals  <- pretty(c(-max(df_plot_data$SIZE), max(df_plot_data$SIZE)))
tick_text  <- prettyNum(abs(tick_vals), digits = 1, big.mark = ",")

plt_popn_proj <- df_plot_data %>%
  plot_ly(
    x = ~if_else(GENDER=="M", SIZE, -SIZE),
    y = ~AGE_BAND_LABEL,
    frame = ~YEAR,
    text = ~LABEL,
    color = ~GENDER,
    colors = c("M" = "#293894", "F" = "#00abbd"),
    type = "bar",
    orientation = "h",
    textposition = "none",
    hovertemplate = "%{text}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    barmode = "relative",
    xaxis = list(
      title = "Population Size",
      tickmode = "array",
      tickvals = tick_vals,
      ticktext = tick_text,
      zeroline = TRUE
    ),
    yaxis = list(
      title = "Age Band",
      categoryorder = "array",
      categoryarray = age_bands_cat
    ),
    legend = list(orientation = "h", x = 0, y = -0.05),
    margin = list(l = 70, r = 30, t = 50, b = 40)
  ) %>%
  animation_opts(frame = 400, easing = "cubic-in-out", redraw = FALSE) %>%
  animation_slider(currentvalue = list(prefix = "Year: "))
```

:::: {.columns}

::: {.column style="width: 30%; font-size: 18px;"}

The population projection at gender and 5 year age band at ICB level^[[ONS Population Projections for sub-Integrated Care Boards by Single Year of Age and Sex](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationprojections/datasets/populationprojectionsforsubintegratedcareboardsbysingleyearofageandsexengland)] is is used to forecast the current population from 2022 to 2047.

Click play to see the annual population projection.

The population of Somerset gets noticeably older each year, for example as of 2025 a quarter (25.4%) of the Somerset population were aged 65 and over, by 2047 that is expected to be 30.1% and this is even more clear for the older age groups, for example as of 2025 around 7.4% of the population were aged 80 and over, by 2047 that is expected to be 12%.

:::

::: {.column width="70%"}

```{r}
#| fig.height: 6
#| out-width: 100%

plt_popn_proj %>% config(responsive = TRUE) 
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

:::: 

## Prevalence {.smaller}

```{r}
#| echo: false
#| warning: false

conditions_short_names <- df_cond$COND
conditions_long_names <- df_cond$COND_LONG
names(conditions_long_names) <- conditions_short_names

df_plot_data <- df_prev %>% 
  pivot_longer(cols = 2:NCOL(.), names_to = "COND", values_to = "PREV") %>%
  mutate(GENDER = str_sub(IDX, 1, 1),
         AGE_BAND = str_sub(IDX, 3),
         AGE_BAND_LABEL = if_else(str_sub(IDX, 3)=="90_PLUS", "90+", gsub("_", "-", AGE_BAND)),
         COND_SHORT = COND, COND_LONG = unname(conditions_long_names[COND]),
         PREV, .keep = "none") %>%
  mutate(LABEL = sprintf("<b>[%s] - %s</b><br>%s %s<br>%.2f%%",
                         COND_SHORT, COND_LONG, 
                         if_else(GENDER=="F", "Female", "Male"), AGE_BAND_LABEL,
                         PREV*100))

chart_title = "Prevalence by 5 Year Age Band for Females"
plt_prev_female <- plot_ly(data = df_plot_data %>% filter(GENDER == "F")) %>%
  add_trace(type = "bar", 
            x = ~AGE_BAND_LABEL, 
            y = ~PREV, 
            color = ~COND_SHORT,
            text = ~LABEL,
            textposition = "none",
            hovertemplate = "%{text}<extra></extra>") %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    font = list(size = 9),
    barmode = "group",
    xaxis = list(
      title = "Age Band"
    ),
    yaxis = list(
      title = "Prevalence",
      tickformat = ".1%",
      zeroline = TRUE
    ),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
        itemclick="toggleothers",
        itemdoubleclick="toggleothers"
    )
  )

chart_title = "Prevalence by 5 Year Age Band for Males"
plt_prev_male <- plot_ly(data = df_plot_data %>% filter(GENDER == "M")) %>%
  add_trace(type = "bar", 
            x = ~AGE_BAND_LABEL, 
            y = ~PREV, 
            color = ~COND_SHORT,
            text = ~LABEL,
            textposition = "none",
            hovertemplate = "%{text}<extra></extra>") %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    font = list(size = 9),
    barmode = "group",
    xaxis = list(
      title = "Age Band"
    ),
    yaxis = list(
      title = "Prevalence",
      tickformat = ".1%",
      zeroline = TRUE
    ),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
      itemclick="toggleothers",
      itemdoubleclick="toggleothers"
    )
  )
```

:::: {.rows}

::: {.row style="width: 100%; height: 20%; font-size: medium;"}

From these charts we can see the impact age has on prevalence for each of the conditions and by gender. The left hand chart shows the **female** prevalence then right the **male**. Clicking on the condition in the legend will show the chart for just that condition, click again to reset to all conditions.

:::

::: {.row style="height: 80%;"}

::: {.column style="width: 50%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 6
#| out-width: 100%

plt_prev_female %>% config(responsive = TRUE) 
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```


:::

::: {.column style="width: 50%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 6
#| out-width: 100%

plt_prev_male %>% config(responsive = TRUE) 
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

:::

::::


## Odds Ratios {.smaller}

```{r}
#| echo: false
#| warning: false

mat_names <- levels(as.factor(c(df_or$COND_I, df_or$COND_J)))

mat <- matrix(data = NA, nrow = length(mat_names), ncol = length(mat_names), dimnames = list(mat_names, mat_names))

for(i in 1:NROW(df_or)){
  mat[df_or$COND_I[i], df_or$COND_J[i]] <- df_or$OR[i]
  mat[df_or$COND_J[i], df_or$COND_I[i]] <- df_or$OR[i]
}

df_plot_data <- mat %>%
  as_tibble() %>% 
  mutate(COND_I = mat_names, .before = 1) %>%
  pivot_longer(-COND_I, names_to = "COND_J", values_to = "OR") %>%
  mutate(
    COND_I_LONG = unname(conditions_long_names[COND_I]),
    COND_J_LONG = unname(conditions_long_names[COND_J]),
    COND_I = factor(COND_I, levels = names(conditions_long_names)),
    COND_J = factor(COND_J, levels = rev(names(conditions_long_names)))
  ) %>%
  mutate(LABEL = if_else(is.na(OR), 
                         "", 
                         sprintf("<b>[%s] - %s<br>[%s] - %s</b><br>OR = %.2f",
                         COND_I, COND_I_LONG, COND_J, COND_J_LONG, OR)))

plt_odd_ratios <- ggplot(df_plot_data, aes(x = COND_I, y = COND_J, tooltip = LABEL)) +
  theme_bw(base_size = 10) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Pairwise Odds Ratios for Conditions", x = "Condition A", y = "Condition B") + 
  guides(fill = guide_none()) +
  ggiraph::geom_tile_interactive(aes(fill = log(OR))) +
  geom_text(aes(label = round(OR, 1))) +
  scale_x_discrete(position = "top") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = log(1), na.value = "grey80")

plt_odd_ratios_interactive <- ggiraph::girafe(ggobj = plt_odd_ratios, width_svg = 8)
plt_odd_ratios_interactive <- girafe_options(
  plt_odd_ratios_interactive,
  opts_tooltip(
    css = "background-color: rgba(255,255,255,0.95); color: black; padding: 10px 12px;
        border-radius: 8px; font-size: 13px; font-family: Arial, Helvetica, sans-serif;
        box-shadow: 0px 2px 6px rgba(0,0,0,0.3); border: 1px solid rgba(255,255,255,0.1);"))
```

:::: {.columns}

::: {.column style="width: 30%; font-size: medium"}

The following chart shows the odds ratios between conditions

- An odds ratio of **zero** indicates a perfect negative relationship, i.e. if condition A is present then it is **impossible** that condition B is present.
- Odds ratios of **greater than 0 and less than 1** indicate a negative relationship, i.e. if condition A is present then it is **less** likely that condition B is present, with the strength of the relationship reducing as the value increases.
- An odds ratio of **one** indicates **independence** i.e. the presence of condition A had no impact on the likelihood of condition B being present.
- An odds ratio of **greater than 1** indicates a positive relationship i.e. if condition A is present then it is **more** likely that condition B is present, with the strength of the relationship increasing as the value increases.

:::

::: {.column style="width: 70%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 8
#| out-width: 100%

plt_odd_ratios_interactive %>% config(responsive = TRUE)
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

::::

## Modelled Prevalence

```{r}
#| echo: false
#| warning: false

df_plot_data <- df_sim %>% 
  mutate(across(all_of(setdiff(names(.), c("STRATA", "YEAR", "COUNT"))), .fns = function(x){x*COUNT})) %>%
  group_by(YEAR) %>%
  summarise(across(all_of(setdiff(names(.), c("STRATA", "YEAR", "COUNT"))), .fns = sum)) %>%
  ungroup() %>%
  pivot_longer(cols = all_of(setdiff(names(.), c("YEAR", "SIZE"))), names_to = "COND", values_to = "REGISTER") %>%
  mutate(COND_LONG = unname(conditions_long_names[COND]),
         PREV = REGISTER/SIZE,
         POPN = SIZE,
         LABEL = sprintf("%s<br><b>[%s] - %s</b><br>Prevalence = %.2f%% (%s of %s)",
                         YEAR, COND, COND_LONG, 
                         PREV * 100,
                         prettyNum(REGISTER, digits = 1, big.mark = ","),
                         prettyNum(POPN, digits = 1, big.mark = ","))) %>% 
  select(YEAR, COND, COND_LONG, REGISTER, POPN, PREV, LABEL)

chart_title = "Modelled Prevalence by Year and Condition"

plt_modelled_prev <- plot_ly(
  data = df_plot_data,
  x = ~YEAR,
  y = ~PREV,
  text = ~LABEL,
  color = ~COND,
  type = "scatter",
  mode = "lines",
  textposition = "none",
  hovertemplate = "%{text}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    font = list(size = 9),
    xaxis = list(title = "Year"),
    yaxis = list(
      title = "Prevalence",
      tickformat = ".2%"),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
      itemclick="toggleothers",
      itemdoubleclick="toggleothers"
    )
)
```

:::: {.columns}

::: {.column style="width: 30%; font-size: medium"}

The following chart shows the modelled prevalence by year and condition. 

We can see for the majority of conditions there is a general upward trend for the prevalence, which is expected as there is a forecast increase in the overall age of the population and the prevalence of most conditions increases with age.

Clicking on the condition in the legend will show the chart for just that condition, click again to reset to all conditions.

:::

::: {.column style="width: 70%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 8
#| out-width: 100%

plt_modelled_prev %>% config(responsive = TRUE) 
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

::::

## Modelled Multimorbidity - Segmentation

```{r}
#| echo: false
#| warning: false

max_count <- max(df_sim$COUNT)

df_glenday <- data.frame(COUNT = 0:max_count) %>%
  left_join(
    df_sim %>% 
      filter(YEAR==2025) %>%
      group_by(COUNT) %>% 
      summarise(SIZE = sum(SIZE)) %>% 
      ungroup(),
    by = "COUNT") %>%
  replace_na(replace = list(SIZE = 0)) %>%
  mutate(PCT = SIZE/sum(SIZE),
         CUM_PCT = cumsum(PCT),
         LABEL = sprintf("<b>Condition Count = %d</b><br>Percentage = %.2f%% (%s of %s)<br>Cumulative Percentage = %.2f%%", COUNT, PCT*100, SIZE, sum(SIZE), CUM_PCT*100))

# Calculate distance from Glenday sieve percentages
green_break <- which.min(abs(df_glenday$CUM_PCT - 0.5))
yellow_break <- which.min(abs(df_glenday$CUM_PCT - 0.95))
blue_break <- which.min(abs(df_glenday$CUM_PCT - 0.99))

# Create breaks
glenday_breaks <- c(-1, green_break-1, yellow_break-1, blue_break-1, max_count)

df_sim$SEGMENT <- factor(cut(df_sim$COUNT, breaks = glenday_breaks, labels = c("GREEN", "YELLOW", "BLUE", "RED")), levels = c("GREEN", "YELLOW", "BLUE", "RED"))

df_sim$SEGMENT_LABEL <- cut(df_sim$COUNT, 
                            breaks = glenday_breaks, 
                            labels = c(sprintf("GREEN [<=%d]", green_break-1),
                                       sprintf("YELLOW [>%d AND <=%d]", green_break-1, yellow_break-1),
                                       sprintf("BLUE [>%d AND <=%d]", yellow_break-1, blue_break-1),
                                       sprintf("RED [>%d AND <=%d]", blue_break-1, max_count)))

df_plot_data <- df_sim %>% 
  mutate(COUNT) %>%
  filter(YEAR==2025) %>%
  group_by(YEAR, COUNT, SEGMENT, SEGMENT_LABEL) %>% 
  summarise(SIZE = sum(SIZE), .groups = "keep") %>%
  ungroup() %>%
  mutate(PCT = SIZE/sum(SIZE),
         CUM_PCT = cumsum(PCT),
         LABEL = sprintf("<b>Condition Count = %d<br>%s</b><br>Percentage = %.2f%% (%s of %s)<br>Cumulative Percentage = %.2f%%", 
                         COUNT, SEGMENT_LABEL, PCT*100, 
                         prettyNum(SIZE, digits = 1, big.mark = ","), 
                         prettyNum(sum(SIZE), digits = 1, big.mark = ","), 
                         CUM_PCT*100))

# Create charts - Overall

chart_title = "Modelled Population by Condition Count and Segment"

plt_segmentation_overall <- plot_ly(
  data = df_plot_data,
  x = ~COUNT,
  y = ~SIZE,
  color = ~SEGMENT,
  colors = c("GREEN" = 'forestgreen', "YELLOW" = 'goldenrod', "BLUE" = 'royalblue', "RED" = "darkred"),
  customdata = ~LABEL,
  type = "bar",
  textposition = "none",
  hovertemplate = "%{customdata}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    barmode = "group",
    font = list(size = 9),
    xaxis = list(title = "Condition Count",
                 tickmode = "array",
                 tickvals = (0:max_count)),
    yaxis = list(
      title = "Population Size",
      tickformat = ","),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
      itemclick="toggleothers",
      itemdoubleclick="toggleothers"
    )
)

# Create charts - Glenday

chart_title = "Modelled Population by Segment"

df_plot_data <- df_sim %>% 
  mutate(COUNT) %>%
  filter(YEAR==2025) %>%
  group_by(YEAR, SEGMENT, SEGMENT_LABEL) %>% 
  summarise(SIZE = sum(SIZE), .groups = "keep") %>%
  ungroup() %>%
  mutate(PCT = SIZE/sum(SIZE),
         CUM_PCT = cumsum(PCT),
         LABEL = sprintf("<b>%s</b><br>Percentage = %.2f%% (%s of %s)<br>Cumulative Percentage = %.2f%%", 
                         SEGMENT_LABEL, PCT*100, 
                         prettyNum(SIZE, digits = 1, big.mark = ","), 
                         prettyNum(sum(SIZE), digits = 1, big.mark = ","), 
                         CUM_PCT*100))

plt_segmentation_glenday <- plot_ly(
  data = df_plot_data,
  x = ~SEGMENT,
  y = ~SIZE,
  color = ~SEGMENT_LABEL,
  colors = c('forestgreen', 'goldenrod', 'royalblue', "darkred"),
  customdata = ~LABEL,
  type = "bar",
  textposition = "none",
  hovertemplate = "%{customdata}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    font = list(size = 9),
    xaxis = list(title = "Segment"),
    yaxis = list(
      title = "Population Size",
      tickformat = ","),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
      itemclick="toggleothers",
      itemdoubleclick="toggleothers"
    )
)
```
:::: {.rows}

::: {.row style="height: 20%; font-size: medium"}

The following charts the segmentation by condition and the overall segment sizes.

:::

::: {.row style="height: 80%;"}

::: {.column style="width: 50%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 8
#| out-width: 100%

plt_segmentation_overall %>% config(responsive = TRUE) 
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

::: {.column style="width: 50%;"}

```{r} 
#| echo: false
#| warning: false
#| fig-width: 8
#| out-width: 100%

plt_segmentation_glenday %>% config(responsive = TRUE)
# %>% htmlwidgets::onRender("function(el){setTimeout(() => Plotly.Plots.resize(el), 0);}")
```

:::

:::

::::

## Modelled Multimorbidity - Forecast

```{r}
#| echo: false
#| warning: false

# Create charts - Glenday

chart_title = "Modelled Population Forecast by Segment"

df_plot_data <- df_sim %>% 
  mutate(COUNT) %>%
  group_by(YEAR, SEGMENT, SEGMENT_LABEL) %>% 
  summarise(SIZE = sum(SIZE), .groups = "keep") %>%
  ungroup() %>%
  group_by(YEAR) %>%  
  mutate(PCT = SIZE/sum(SIZE),
         CUM_PCT = cumsum(PCT),
         LABEL = sprintf("<b>%d<br>%s</b><br>Percentage = %.2f%% (%s of %s)", 
                         YEAR, SEGMENT_LABEL, PCT*100, 
                         prettyNum(SIZE, digits = 1, big.mark = ","), 
                         prettyNum(sum(SIZE), digits = 1, big.mark = ",")))

plt_segmentation_forecast <- plot_ly(
  data = df_plot_data,
  x = ~YEAR,
  y = ~PCT,
  color = ~SEGMENT_LABEL,
  colors = c('forestgreen', 'goldenrod', 'royalblue', "darkred"),
  customdata = ~LABEL,
  type = "scatter",
  mode = "line",
  textposition = "none",
  hovertemplate = "%{customdata}<extra></extra>"
  ) %>%
  layout(
    autosize = TRUE,
    title = chart_title,
    font = list(size = 9),
    xaxis = list(title = "Year"),
    yaxis = list(
      title = "% of Population",
      tickformat = ".2%"),
    margin = list(l = 70, r = 30, t = 50, b = 40),
    legend=list(
      itemclick="toggleothers",
      itemdoubleclick="toggleothers"
    )
)
```

:::: {.columns}

::: {.column style="width: 30%; font-size: large"}

The chart shown here shows the forecast for each segment for each year.

We can see the decreasing proportion of green segment and increasing proportion of yellow, blue and red segments indicating increased demand for health services.
:::

::: {.column style="width: 70%;"}

```{r}
#| fig-height: 6
#| out-width: 100%

plt_segmentation_forecast
```

:::

::::
